\mytitle

### 2020-04-14 paper.tex
Edited to end of Sect 2.1

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are naturally captured using quadratic
constraints and objective functions. However, with a few exceptions (e.g., \cite{turner2014new}), exact optimization
approaches that have been proposed for the WFLO
problem~\cite{Zhang2014,kuo2016wind} \todo{[Add more references here]}
represent the quadratic relations using linear models
that approximate the true dependencies.

@@

We start this section
by presenting the
physical model 
of a wind farm, which we use throughout the paper.
Our wind farm model is based on
the common notation and description 
presented in~\cite{Zhang2014}.  
Subsequently, we provide a literature review of 
existing mathematical
optimization methods for solving WFLO, focusing
on their similarities and differences
compared to our approach.

\subsection{The Wind Farm Model}
\begin{figure}[t]
	\centering
	\includegraphics[scale = 0.9]{field_model.pdf}

@@

\paragraph{Wind Farm as a Grid} We consider a square shaped wind farm that we model 
as an $n\times n$ grid
with $n$ being the number of cells on each axis. 
The cells have an area of $c^2$ (in squared meters), i.e.,
if $c=100$ meters and $n=20$, the farm area will be $2km\times 2km$.
In this work, we assume that the terrain is flat and that 
we have no constraints related to the terrain.
Furthermore, we do not account for the influence of noise
on the wind farm's surroundings. The wind farm literature treats the more complex
farm models that capture non-flat terrain~\cite{song2015lazy,kuo2016wind}, multi-typed turbines~\cite{feng2017design}, noise effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex objectives (including turbine installation  and maintenance costs)~\cite{lackner2007analytical}.
In this work, we focus on the novel quadratic approach for WFLO, and hence
consider the most abstract wind farm model. Yet, our model can be extended,
in the future, to include 
additional wind farm settings. \todo{perhaps discuss in future work}

Returning to the grid model, we assume that the number of homogeneous turbines to be placed is known 
in advance; we denote the number of turbines by $m$. To place these $m$ turbines, 
we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
In what follows, we explain the two effects using a toy example of placing a single turbine on an 
$8\times8$ grid (see Figure~\ref{fig:field_model}).

\paragraph{Proximity Constraints} Turbines must be placed at least \emph{five rotor diameters} apart. Depending on the cell size, 
	we shall formulate the corresponding constraint as part of the optimization model. For example, assume that a turbine is placed
	in the center of a cell. Then, given a cell size of $c = 100$ meters and a rotor diameter of $40$ meters, 
	we must ensure that two turbines will not placed inside two adjacent cells. In Figure~\ref{fig:field_model}, 
	the red area around the turbine represents the forbidden cells.

\paragraph{Wakes} Turbines influence each other through interference effects that are referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}. 
	 Wakes cause reduction of the effective power produced by a turbine due to upstream turbines that change
	 its airflow dynamics. In Figure~\ref{fig:field_model}, the wind is assumed to be horizontal. Therefore, the wake effect due
	 to placing the turbine expands as shown in the diagram: it will influence only turbines that will be placed in the orange cells.
	 In other words, if a turbine is placed in the orange area, its effective
	 power is reduced due to the upstream turbine that we placed in the $8\times8$ farm. 
	 Moreover, wakes can occur between multiple turbines, and hence
	 multiple wakes must be taken into account for each location. 
	 The reduction of power due to the wake effect is determined by three parameters: (1) the distance between the upstream
	 turbine and the cell for which we want to compute the effect (denoted by $x$ in Figure~\ref{fig:field_model}), (2) the radius of the cone's opening 
	 (denoted by $r$ in Figure~\ref{fig:field_model}), and (3) 
	 the superposition of several wind regimes (a wind regime is a combination of free speed and directions).
	 The wake effect can be pre-computed for each pair of cells 
	 since we know: the distance ($x$), the turbine specifications and terrain conditions (that jointly dictate $r$), and  
	 probabilistic behavior of wind regimes~\cite{Zhang2014}.

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes. Every regime $d\in \mathcal{D}$ is a unique combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind blowing at a free velocity of $12km/h$. The
probability of this regime is $0.008$. Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
Note that turbines can adjust their direction according to the current wind 
regime.

Given that we must consider wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (second component of regime $d$), the set of turbines $\mathcal{U}_{id}$ that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake 
reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. Furthermore, the set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid w.r.t. to the current wind regime $d \in \mathcal{D}$ and computing which turbine $j$ is located upstream to turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}): \begin{equation} E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d. \end{equation} This is also known to be the most accurate analytical total power expression that accounts for multiple wakes. However, the expression is not amenable for optimization, and hence, a simplified representation for the total expected energy that considers wakes is required.

An alternative 
form of the total expected power in the presence of multiple wakes is the linear superposition (LS) expression: \begin{equation} \label{eq:ls}
E_{LS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \Bigg(\frac{1}{3}u_{id\infty}^3 -\sum_{j\in\mathcal{U}_{id}} \frac{1}{3}(u_{id\infty}^3 - u_{ijd}^3)\Bigg).
\end{equation} Note that $E_{LS}$ has the same parameters as $E_{SS}$; the difference is in the function that ties these parameters together. 
The LS model is known to be less accurate compared to SS, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.

>>>>

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are naturally captured using quadratic
constraints and objective functions. However, with a few exceptions (e.g., \cite{turner2014new}), exact optimization
approaches that have been proposed for the WFLO
problem~\cite{Zhang2014,kuo2016wind} \todo{[JCB: Add more references here]}
represent the quadratic relations using linear models
that approximate the true dependencies.

@@

We start this section by presenting the physical model of a wind farm
used throughout the paper.  Our wind farm model is based on the common
notation and description presented in~\cite{Zhang2014}.  Subsequently,
we provide a literature review of existing mathematical optimization
methods for solving WFLO, focusing on their similarities and
differences with our approach.

\subsection{The Wind Farm Model}

%\paragraph{Wind Farm as a Grid}
We consider a square-shaped wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in squared meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  In this work, we
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical}.  In this work, we focus on the
novel quadratic approach for WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional wind farm settings. \todo{perhaps
  discuss in future work}

We assume that the number of identical turbines to be placed is known 
in advance and denote the number of turbines by $m$. To place these $m$ turbines, 
we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
In what follows, we explain the two effects using a toy example of placing a single turbine on an 
$8\times8$ grid (see Figure~\ref{fig:field_model}).

\begin{figure}[t]
	\centering
	\includegraphics[scale = 0.9]{field_model.pdf}

@@

\paragraph{Proximity Constraints}
Turbines must be placed at least \emph{five rotor diameters}
apart. Depending on the cell size, we shall formulate the
corresponding constraint as part of the optimization model. For
example, assume that a turbine is placed in the center of a
cell. Then, given a cell size of $c = 100$ meters and a rotor diameter
of $40$ meters, we must ensure that two turbines are not placed in
two adjacent cells. In Figure~\ref{fig:field_model}, the red area
around the turbine represents the forbidden cells.

\paragraph{Wakes}
Turbines influence each other through interference effects that are
referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}.
Wakes cause reduction of the effective power produced by a turbine due
to upstream turbines that change the airflow dynamics. In
Figure~\ref{fig:field_model}, the wind is assumed to be
blowing from left to right. The wake effect due to placing the turbine
expands as shown in the diagram.
%: it will influence only turbines that
%are placed in the orange cells.  In other words,
If a turbine is placed in the orange area, its effective power is
reduced due to the upstream turbine.
% that we placed in the $8\times8$ farm.
Moreover, a given turbine may be placed downstream from multiple turbines and hence
multiple wakes must be taken into account for each location.  The
reduction of power due to the wake effect is determined by three
parameters: (1) the distance between the upstream turbine and the cell
for which we want to compute the effect (denoted by $x$ in
Figure~\ref{fig:field_model}), (2) the radius of the cone's opening
(denoted by $r$ in Figure~\ref{fig:field_model}), and (3) the
superposition of several wind regimes (a wind regime is a combination
of free speed and directions).  The wake effect can be pre-computed
for each pair of cells since we know: the distance ($x$), the turbine
specifications and terrain conditions (that jointly dictate $r$), and
probabilistic behavior of wind regimes~\cite{Zhang2014}.

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a unique \todo{[JCB: Mathematically, I don't think it matters if the regime is unique. So delete the word?]} combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-eastern wind blowing at a free velocity of $12\mbox{ km/h}$. The
probability of this regime is $0.008$.

Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. 
Note that turbines can adjust their direction according to the current wind 
regime.

Given the interference of the wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of $d$.
The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. The set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid w.r.t.\ to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}):
\begin{equation}
  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
\end{equation}

  Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes.\todo{[JCB: REF?]} However, the expression is not amenable to optimization \todo{[JCB: because?]}. Hence, a simplified representation for the total expected energy that considers wakes is required.

An alternative 
form of the total expected power in the presence of multiple wakes is the linear superposition (LS) expression:

\begin{equation} \label{eq:ls}
E_{LS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \Bigg(\frac{1}{3}u_{id\infty}^3 -\sum_{j\in\mathcal{U}_{id}} \frac{1}{3}(u_{id\infty}^3 - u_{ijd}^3)\Bigg).
\end{equation}

Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
The LS model is less accurate compared to SS,\todo{[JCB: REF?]} yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint. \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic. Sounds like it is not natural and we choose it due to its optimization properties.]}
